### 2018-01-25 paper/ms.tex
words

\documentclass[rnaas]{aastex62}

\pdfoutput=1

\usepackage{lmodern}
\usepackage{microtype}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{multirow}
\usepackage{graphicx}
\bibliographystyle{aasjournal}

% ------------------ %
% end of AASTeX mods %
% ------------------ %

% Projects:
\newcommand{\project}[1]{\textsf{#1}}
\newcommand{\kepler}{\project{Kepler}}
\newcommand{\ktwo}{\project{K2}}

% references to text content
\newcommand{\documentname}{\textsl{Note}}
\newcommand{\figureref}[1]{\ref{fig:#1}}
\newcommand{\Figure}[1]{Figure~\figureref{#1}}
\newcommand{\figurelabel}[1]{\label{fig:#1}}
\renewcommand{\eqref}[1]{\ref{eq:#1}}
\newcommand{\Eq}[1]{Equation~(\eqref{#1})}
\newcommand{\eq}[1]{\Eq{#1}}
\newcommand{\eqalt}[1]{Equation~\eqref{#1}}
\newcommand{\eqlabel}[1]{\label{eq:#1}}

% TODOs
\newcommand{\todo}[3]{{\color{#2}\emph{#1}: #3}}
\newcommand{\dfmtodo}[1]{\todo{DFM}{red}{#1}}
\newcommand{\alltodo}[1]{\todo{TEAM}{red}{#1}}
\newcommand{\citeme}{{\color{red}(citation needed)}}

% math
\newcommand{\T}{\ensuremath{\mathrm{T}}}
\newcommand{\dd}{\ensuremath{ \mathrm{d}}}
\newcommand{\unit}[1]{{\ensuremath{ \mathrm{#1}}}}
\newcommand{\bvec}[1]{{\ensuremath{\boldsymbol{#1}}}}
\newcommand{\Gaussian}[3]{\ensuremath{\frac{1}{|2\pi #2|^\frac{1}{2}}
            \exp\left[ -\frac{1}{2}#1^\top #2^{-1} #1 \right]}}

% VECTORS AND MATRICES USED IN THIS PAPER
\newcommand{\Normal}{\ensuremath{\mathcal{N}}}
\newcommand{\mA}{\ensuremath{\bvec{A}}}
\newcommand{\mC}{\ensuremath{\bvec{C}}}
\newcommand{\mS}{\ensuremath{\bvec{\Sigma}}}
\newcommand{\mL}{\ensuremath{\bvec{\Lambda}}}
\newcommand{\vw}{\ensuremath{\bvec{w}}}
\newcommand{\vy}{\ensuremath{\bvec{y}}}
\newcommand{\vt}{\ensuremath{\bvec{\theta}}}
\newcommand{\vm}{\ensuremath{\bvec{\mu}(\bvec{\theta})}}
\newcommand{\vre}{\ensuremath{\bvec{r}}}
\newcommand{\vh}{\ensuremath{\bvec{h}}}
\newcommand{\vk}{\ensuremath{\bvec{k}}}

% typography obsessions
\setlength{\parindent}{3.0ex}

\begin{document}\raggedbottom\sloppy\sloppypar\frenchspacing

\title{%
    Scalable gradients of Gaussian Process models with celerite
}

\author[0000-0002-9328-5652]{Daniel Foreman-Mackey}
\affil{Center for Computational Astrophysics, Flatiron Institute, New York, NY}

\keywords{%
methods: data analysis ---
methods: statistical
}

\section{Abstract}

This research note presents a derivation and implementation of reverse-mode
gradients of the \emph{celerite} algorithm for scalable inference with
Gaussian Process (GP) models.
These gradients can be integrated into existing automatic differentiation
frameworks to provide a scalable and computationally efficient method for
evaluating the gradients of the GP likelihood with respect to the kernel
hyperparameters.
The algorithm derived in this note uses less memory and is more efficient than
versions using automatic differentiation and the computational cost scales
linearly with the number of data points.

\section{Introduction}

Gaussian Processes \citep[GPs][]{Rasmussen:2006} are a class of models used
extensively in the astrophysical literature to model stochastic processes.
The applications are broad-ranging and some example include the time domain
variability of astronomical sources \citep{Brewer:2009, Kelly:2014,
Haywood:2014, Rajpaul:2015, Foreman-Mackey:2017}, data-driven models of light
curves or stellar spectra \citep{Wang:2012, Luger:2016, Czekala:2017}, and the
cosmic microwave background \citep{Bond:1987,Wandelt:2003}.
In all of these applications, the calculation and optimization of the GP
marginalized likelihood function (here we follow the notation of
\citealt{Foreman-Mackey:2017})
\begin{eqnarray}\eqlabel{loglike}
\log \mathcal{L}(\bvec{\theta},\,\bvec{\alpha}) &=&
    -\frac{1}{2}\,\left[\bvec{y} - \bvec{\mu}_\bvec{\theta}\right]^\T\,
        {K_\bvec{\alpha}}^{-1}\,\left[\bvec{y}-\bvec{\mu}_\bvec{\theta}\right]
    -\frac{1}{2}\,\log\det K_\bvec{\alpha} + \mathrm{constant}
\end{eqnarray}
is of central importance and this can often be the computational bottleneck.
The details of this model are omitted here (more details can be found in
\citealt{Rasmussen:2006} and \citealt{Foreman-Mackey:2017}), but the key point
is that, for a dataset with $N$ data points, every evaluation of a GP model
requires computation of the inverse\footnote{Or, more specifically, a linear
operator that can left multiply a vector or a matrix by the inverse of this
matrix.} and log-determinant of the $N \times N$ covariance matrix.
The computational cost of these operations scales as $\mathcal{O}(N^3)$ in the
general case, but the \emph{celerite} method was recently introduced in the
astronomical literature to compute the GP likelihood for a class of
one-dimensional models with $\mathcal{O}(N)$ scaling
\citep{Foreman-Mackey:2017}.

The details of the \emph{celerite} method can be found in
\citet{Foreman-Mackey:2017} and I will not repeat them here.
The only difference in notation here is that all the matrices to which I refer
are the ``pre-conditioned'' matrices that are indicated with a tilde by
\citet{Foreman-Mackey:2017}, but the tilde is not included here for
simplicity.
I will also use the symbol $P$ to mean the $N-1 \times J$ pre-conditioning
matrix that was called $\phi$ by \citet{Foreman-Mackey:2017}.
The Cholesky factorization algorithm derived by \citet[][their
Equation~46]{Foreman-Mackey:2017} is given is listed below in the following
algorithm:

\medskip
\begin{minipage}{\linewidth}
\textbf{function} \texttt{celerite\_factor}($U$, $P$, $\bvec{d}$, $W$) \\
\hspace*{2em}\textsf{\# At input $\bvec{d} = \bvec{a}$ and $W = V$} \\
\hspace*{2em}$S \gets$ \texttt{zeros}($J$, $J$) \\
    \hspace*{2em}$\bvec{w}_1 \gets \bvec{w}_1 / d_{1}$ \\
\hspace*{2em}\textbf{for} $n = 2,\ldots,N$:\\
\hspace*{2em}\hspace*{2em}$S \gets \texttt{diag}(\bvec{p}_{n-1})\,[
    S + d_{n-1}\,{\bvec{w}_{n-1}}^\T\,{\bvec{w}_{n-1}}
]\,\texttt{diag}(\bvec{p}_{n-1})$ \\
\hspace*{2em}\hspace*{2em}$d_{n} \gets d_{n} - \bvec{u}_n\,S\,{\bvec{u}_n}^\T$\\
\hspace*{2em}\hspace*{2em}$\bvec{w}_n \gets \left[\bvec{w}_n -
    \bvec{u}_n\,S \right] / d_{n}$\\
    \hspace*{2em}\textbf{return} $\bvec{d}$, $W$, $S$
\end{minipage}
\medskip

\noindent In this algorithm, the \texttt{zeros}$(J,\,K)$ function creates a $J
\times K$ matrix of zeros, the \texttt{diag} function creates a diagonal
matrix from a vector, and $\bvec{x}_n$ indicates a row vector made from the
$n$-th row of the matrix $X$.
Using this factorization, the log-determinant of $K$ is
\begin{eqnarray}
\log \det K &=& \sum_{n=1}^N \log d_{n} \quad.
\end{eqnarray}
Similarly, \citet{Foreman-Mackey:2017} derived a $\mathcal{O}(N\,J)$ algorithm
to apply the inverse of $K$ as follows (Equations~47 and~48 in
\citealt{Foreman-Mackey:2017}):

\medskip
\begin{minipage}{\linewidth}
\textbf{function} \texttt{celerite\_solve}($U$, $P$, $\bvec{d}$, $W$, $Z$) \\
\hspace*{2em}\textsf{\# At input $Z = Y$} \\
\hspace*{2em}$F \gets$ \texttt{zeros}($J$, $N_\mathrm{rhs}$) \\
\hspace*{2em}\textbf{for} $n = 2,\ldots,N$:\\
\hspace*{2em}\hspace*{2em}$F \gets \texttt{diag}(\bvec{p}_{n-1})\,[F +
    {\bvec{w}_{n-1}}^\T\,\bvec{z}_{n-1}]$\\
\hspace*{2em}\hspace*{2em}$\bvec{z}_n \gets \bvec{z}_n - \bvec{u}_n\,F$\\
\hspace*{2em}\textbf{for} $n = 1,\ldots,N$:\\
\hspace*{2em}\hspace*{2em}$\bvec{z}_n \gets \bvec{z}_n / d_{n}$\\
\hspace*{2em}$G \gets$ \texttt{zeros}($J$, $N_\mathrm{rhs}$) \\
\hspace*{2em}\textbf{for} $n = N-1,\ldots,1$:\\
\hspace*{2em}\hspace*{2em}$G \gets \texttt{diag}(\bvec{p}_{n})\,[G +
    {\bvec{u}_{n+1}}^\T\,\bvec{z}_{n+1}]$\\
    \hspace*{2em}\hspace*{2em}$\bvec{z}_n \gets \bvec{z}_n - \bvec{w}_n\,G$\\
\hspace*{2em}\textbf{return} $Z$, $F$, $G$
\end{minipage}
\medskip

\section{Gradients of GP models using celerite}

The standard method of computing gradients of \Eq{loglike} with respect to the
parameters $\bvec{\theta}$ and $\bvec{\alpha}$ is the equation
\citep{Rasmussen:2006}
\begin{eqnarray}\eqlabel{naive-grad}
\frac{\dd \log \mathcal{L}}{\dd \alpha_k} &=&
    \frac{1}{2}\,\mathrm{Tr}\left[
        \left[
        \bvec{\tilde{r}}\,\bvec{\tilde{r}}^\T - {K_\bvec{\alpha}^{-1}}
        \right]
        \,\frac{\dd K}{\dd \alpha_k}
    \right]
\end{eqnarray}
where
\begin{eqnarray}
    \bvec{\tilde{r}} &=&
        {K_\bvec{\alpha}}^{-1}\,\left[\bvec{y}-\bvec{\mu}_\bvec{\theta}\right]
    \quad.
\end{eqnarray}
Similar equations exist for the parameters $\bvec{\theta}$.
Even with a scalable method of applying ${K_\bvec{\alpha}}^{-1}$, the
computational cost of \Eq{naive-grad} scales as $\mathcal{O}(N^2)$.
This scaling is prohibitive when applying the \emph{celerite} method to large
datasets and I have not found a simple analytic method of improving this
scaling for semi-separable matrices.
However, it was recently demonstrated that substantial computational gains can
be made by directly differentiating Cholesky factorization algorithms even in
the general case \citep{Murray:2016}.

Following this reasoning and using the notation from an excellent review of
matrix gradients \citep{Giles:2008}, I present the reverse-mode gradients of
the \emph{celerite} method.
While not yet popular within astrophysics, ``reverse-mode accumulation'' of
gradients (also known as ``backpropagation'') has revolutionalized the field
of machine learning over the past few decades (TODO ADD CITATION).
The review \citep{Giles:2008} provides a thorough overview of these methods
and the interested reader is directed to that discussion for details and for
an explanation of the notation.

Using the notation from \citet{Giles:2008} and performing some tedious
algebra, the reverse-mode implementation of the \texttt{celerite\_factor}
function is found to be:

\medskip
\begin{minipage}{\linewidth}
\textbf{function} \texttt{celerite\_factor\_rev}($U$, $P$, $\bvec{d}$, $W$,
    $S$, $\bar{S}$, $\bar{\bvec{a}}$, $\bar{V}$) \\
\hspace*{2em}\textsf{\# At input $\bar{\bvec{a}} = \bar{\bvec{d}}$ and
    $\bar{V} = \bar{W}$}\\
    \hspace*{2em}$\bar{\bvec{v}}_N \gets \bar{\bvec{v}}_N / d_N$\\
\hspace*{2em}\textbf{for} $n = N,\ldots,2$:\\
\hspace*{2em}\hspace*{2em}$\bar{a}_n \gets \bar{a}_n -
    \bvec{w}_n\,{\bar{\bvec{v}}_n}^\T$\\
\hspace*{2em}\hspace*{2em}$\bar{\bvec{u}}_n \gets - [\bar{\bvec{v}}_n +
    2\,\bar{a}_n\,\bvec{u}_n]\,S$\\
\hspace*{2em}\hspace*{2em}$\bar{S} \gets \bar{S} -
    {\bvec{u}_n}^\T\,[\bar{\bvec{v}}_n + \bar{a}_n\,\bvec{u}_n]$\\
\hspace*{2em}\hspace*{2em}$\bar{\bvec{p}}_{n-1} \gets \mathrm{diag}(\bar{S}\,S
    \,\mathrm{diag}(\bvec{p}_{n-1})^{-1} + \mathrm{diag}(\bvec{p}_{n-1})^{-1}\,
    S\,\bar{S})$\\
\hspace*{2em}\hspace*{2em}$\bar{S} \gets \mathrm{diag}(\bvec{p}_{n-1})\,
    \bar{S}\,\mathrm{diag}(\bvec{p}_{n-1})$\\
\hspace*{2em}\hspace*{2em}$\bar{d}_{n-1} \gets \bar{d}_{n-1} +
    \bvec{w}_{n-1}\,\bar{S}\,{\bvec{w}_{n-1}}^\T$\\
\hspace*{2em}\hspace*{2em}$\bar{\bvec{v}}_{n-1} \gets \bar{\bvec{v}}_{n-1}
    / d_{n-1} + \bvec{w}_{n-1}\,[\bar{S} + \bar{S}^\T]$\\
\hspace*{2em}\hspace*{2em}$S \gets \mathrm{diag}(\bvec{p}_{n-1})^{-1}\,S\,
    \mathrm{diag}(\bvec{p}_{n-1})^{-1}
    - d_{n-1}\,{\bvec{w}_{n-1}}^\T\,\bvec{w}_{n-1}$\\
\hspace*{2em}$\bar{a}_1 \gets \bar{a}_1 -
    \bar{\bvec{v}}_1\,{\bvec{w}_1}^\T$\\
\hspace*{2em}\textbf{return} $\bar{U}$, $\bar{P}$, $\bar{\bvec{a}}$, $\bar{V}$
\end{minipage}
\medskip

\noindent Similarly, the reverse-mode function corresponding to
\texttt{celerite\_solve} is:

\medskip
\begin{minipage}{\linewidth}
\textbf{function} \texttt{celerite\_solve\_rev}($U$, $P$, $\bvec{d}$, $W$,
    $Z$, $F$, $G$, $\bar{F}$, $\bar{G}$, $\bar{Y}$) \\
\hspace*{2em}\textsf{\# At input $\bar{Y} = \bar{Z}$}\\
\hspace*{2em}\textbf{for} $n = 1,\ldots,N-1$:\\
\hspace*{2em}\hspace*{2em}$\bar{\bvec{w}}_n \gets
    - \bar{\bvec{y}}_n\,G^\T$\\
\hspace*{2em}\hspace*{2em}$\bar{G} \gets \bar{G} - {\bvec{w}_n}^\T \,
    \bar{\bvec{y}}_n$\\
\hspace*{2em}\hspace*{2em}$\bvec{z}_n \gets \bvec{z}_n + \bvec{w}_n\,G$\\
\hspace*{2em}\hspace*{2em}$G \gets \mathrm{diag}(\bvec{p}_n)^{-1}\,G$\\
\hspace*{2em}\hspace*{2em}$\bar{\bvec{p}}_n \gets \mathrm{diag}(
    \bar{G}\,G^\T)$\\
\hspace*{2em}\hspace*{2em}$\bar{G} \gets
    \mathrm{diag}(\bvec{p}_n)\,\bar{G}$\\
\hspace*{2em}\hspace*{2em}$G \gets G - {\bvec{u}_{n+1}}^\T\,\bvec{z}_{n+1}$\\
\hspace*{2em}\hspace*{2em}$\bar{\bvec{u}}_{n+1} \gets
    \bvec{z}_{n+1}\,\bar{G}^\T$\\
\hspace*{2em}\hspace*{2em}$\bar{\bvec{y}}_{n+1} \gets
    \bvec{u}_{n+1}\,\bar{G}$\\
\hspace*{2em}\textbf{for} $n = 1,\ldots,N$:\\
\hspace*{2em}\hspace*{2em}$\bar{\bvec{y}}_n \gets \bar{\bvec{y}}_n/d_n$\\
\hspace*{2em}\hspace*{2em}$\bar{d}_n \gets -
    \bvec{z}_n\,{\bar{\bvec{y}}_n}^\T$\\
\hspace*{2em}\textbf{for} $n = N,\ldots,2$:\\
\hspace*{2em}\hspace*{2em}$\bar{\bvec{u}}_n \gets \bar{\bvec{u}}_n
    - \bar{\bvec{y}}_n\,F^\T$\\
\hspace*{2em}\hspace*{2em}$\bar{F} \gets \bar{F} - {\bvec{u}_n}^\T \,
    \bar{\bvec{y}}_n$\\
\hspace*{2em}\hspace*{2em}$F \gets \mathrm{diag}(\bvec{p}_{n-1})^{-1}\,F$\\
\hspace*{2em}\hspace*{2em}$\bar{\bvec{p}}_{n-1} \gets \bar{\bvec{p}}_{n-1} +
    \mathrm{diag}(\bar{F}\,F^\T)$\\
\hspace*{2em}\hspace*{2em}$\bar{F} \gets
    \mathrm{diag}(\bvec{p}_{n-1})\,\bar{F}$\\
\hspace*{2em}\hspace*{2em}$F \gets F - {\bvec{w}_{n-1}}^\T\,\bvec{z}_{n-1}$\\
\hspace*{2em}\hspace*{2em}$\bar{\bvec{w}}_{n-1} \gets \bar{\bvec{w}}_{n-1} +
    \bvec{z}_{n-1}\,\bar{F}^\T$\\
\hspace*{2em}\hspace*{2em}$\bar{\bvec{y}}_{n-1} \gets \bar{\bvec{y}}_{n-1} +
    \bvec{w}_{n-1}\,\bar{F}$\\
\hspace*{2em}\textbf{return} $\bar{U}$, $\bar{P}$, $\bar{\bvec{d}}$,
    $\bar{W}$, $\bar{Y}$
\end{minipage}
\medskip

\noindent A reference C++ implementation of this algorithm can be found online
TODO ADD ZENODO LINK and \Figure{figure} shows the performance of this
implementation.

\section{Discussion}

This research note presents the algorithms needed to efficiently compute
gradients of GP models applied to large datasets using the \emph{celerite}
method.
These developments increase the performance of inference methods based on
\emph{celerite} and improve the convergence properties of non-linear
optimization routines.
Furthermore, the derivation of reverse-mode accumulation algorithms for
\emph{celerite} allow its integration into popular model building and
automatic differentiation libraries like Stan \citep{Carpenter:2015},
TensorFlow \citep{Abadi:2016}, and others.

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.9\textwidth]{figure.pdf}
\caption{%
    Scaling.
\figurelabel{figure}}
\end{center}
\end{figure}

%\acknowledgements
%It is a pleasure to thank
%  Patrick Cooper,
%  Boris Leistedt,
%  Bernhard Sch\"olkopf, and
%  Dun Wang
%for helping us understand all of this.

\bibliography{celerite}

\end{document}